---
title: "SomaHRD v1.2.8 verification (INC9891)"
author: "Joe Shaw (CS20980)"
format: html
date: today
date-format: "YYYY-MM-DD"
---
```{r}
#| label: packages
#| include: FALSE

library(here)
library(readxl)

source(here("functions/hrd_functions.R"))

data_filepath <- config::get(value = "data_filepath")

options(digits = 16)

```

# Verification of SomaHRD v1.2.8

10 samples were run on SomaHRD v1.2.8 on the SeqOne portal and the csv file 
outputs were downloaded to this file location:

`r paste0(data_filepath, "live_service/INC9891/")`

The results from these files were compared to the csv file outputs for the 
same samples tested using SomaHRD v1.2.7 during INC9110, which are saved here:

`r paste0(data_filepath, "live_service/INC9110/")`

## Overall Results

The results are shown in @tbl-version-comparison.

```{r}
#| label: tbl-version-comparison
#| tbl-cap: "SomaHRD v1.2.7 vs 1.2.8 results"
#| echo: FALSE

v1_2_8_files <- list.files(path = paste0(
  data_filepath,
  "live_service/INC9891/"),
  full.names = TRUE,
  pattern = "*.csv")

v1_2_8_data <- v1_2_8_files |> 
  map(\(v1_2_8_files) read_seqone_csv(file = v1_2_8_files)) |> 
  list_rbind() |> 
  mutate(semantic_version = "v1.2.8")

v1_2_7_files <- list.files(path = paste0(
  data_filepath,
  "live_service/INC9110/20240412_reanalysis/"),
  full.names = TRUE,
  pattern = "*.csv")

v1_2_7_data <- v1_2_7_files |> 
  map(\(v1_2_7_files) read_seqone_csv(file = v1_2_7_files)) |> 
  list_rbind() |> 
  mutate(semantic_version = "v1.2.7") |> 
  filter(sample %in% v1_2_8_data$sample)

both_versions <- rbind(v1_2_7_data, v1_2_8_data) |> 
  arrange(sample, analysis_date) |> 
  select(-somahrd_version) |> 
  pivot_wider(id_cols = c(sample),
              names_from = semantic_version,
              values_from = -c(sample, semantic_version))

knitr::kable(both_versions) 

```

All outputs are equal except for the low_tumour_fraction output when compared 
to 16 decimal places.

```{r}
#| label: check-v-1-2-8
#| include: TRUE

identical(both_versions$LGA_v1.2.7, both_versions$LGA_v1.2.8)

identical(both_versions$LPC_v1.2.7, both_versions$LPC_v1.2.8)

identical(both_versions$score_v1.2.7, both_versions$score_v1.2.8)

identical(both_versions$status_v1.2.7, both_versions$status_v1.2.8)

identical(both_versions$brca_status_v1.2.7, both_versions$brca_status_v1.2.8)

identical(both_versions$brca_mutation_v1.2.7, both_versions$brca_mutation_v1.2.8)

identical(both_versions$ccne1_cn_v1.2.7, both_versions$ccne1_cn_v1.2.8)

identical(both_versions$rad51b_cn_v1.2.7, both_versions$rad51b_cn_v1.2.8)

identical(both_versions$coverage_v1.2.7, both_versions$coverage_v1.2.8)

identical(both_versions$pct_mapped_reads_v1.2.7, both_versions$pct_mapped_reads_v1.2.8)

identical(both_versions$pct_tum_cell_v1.2.7, both_versions$pct_tum_cell_v1.2.8)

identical(both_versions$gi_confidence_v1.2.7, both_versions$gi_confidence_v1.2.8)

identical(both_versions$low_tumor_fraction_v1.2.7, both_versions$low_tumor_fraction_v1.2.8)

```

## Low Tumour Fraction

```{r}
#| label: tumor-fraction-difference
#| include: FALSE

low_tumor_diff <- both_versions |> 
  select(sample, low_tumor_fraction_v1.2.7, low_tumor_fraction_v1.2.8) |> 
  mutate(diff = abs(low_tumor_fraction_v1.2.7 - low_tumor_fraction_v1.2.8))

```

The difference in low tumour fraction calculation varied from 
`r min(low_tumor_diff$diff)` to `r max(low_tumor_diff$diff)`.

```{r}
#| label: tbl-tumor-fraction-difference
#| tbl-cap: "Low tumour fraction difference"
#| echo: FALSE

knitr::kable(low_tumor_diff)

```

# Verification of SomaHRD v1.2.12

On 2024-12-31 SeqOne alerted us that they were updating SomaHRD from v1.2.8 to
v1.2.12.

They had performed a comparison analysis of the same 10 samples, and sent us 
the results: "2024-12-31_Manchester SomaHRD v1.2 patch updates - 
Non-regression testing.xlsx"

```{r}
#| include: FALSE

seqone_filename <- "2024-12-31_Manchester SomaHRD v1.2 patch updates - Non-regression testing.xlsx"

seqone_excel_tab1 <- read_excel(path = paste0(data_filepath,
                                         "live_service/INC9891/",
                                        seqone_filename),
                           sheet = "1.2.7 -> 1.2.8",
                           col_types = 
                             c("text", "date", "text", 
                               "numeric", "numeric", "numeric", 
                               "text", "text", "text",
                               "numeric", "numeric", "numeric", 
                               "numeric", "numeric", "numeric", 
                               "numeric")) |> 
  janitor::clean_names()

seqone_excel_tab2 <- read_excel(path = paste0(data_filepath,
                                         "live_service/INC9891/",
                                        seqone_filename),
                           sheet = "1.2.8 -> 1.2.12",
                           col_types = 
                             c("text", "date", "text", 
                               "numeric", "numeric", "numeric", 
                               "text", "text", "text",
                               "numeric", "numeric", "numeric", 
                               "numeric", "numeric", "numeric", 
                               "numeric")) |> 
  janitor::clean_names()

seqone_tab1_wide <- seqone_excel_tab1 |> 
  pivot_wider(id_cols = sample,
              names_from = somahrd_version,
              values_from = -c(sample, somahrd_version))

seqone_tab2_wide <- seqone_excel_tab2 |> 
  pivot_wider(id_cols = sample,
              names_from = somahrd_version,
              values_from = -c(sample, somahrd_version))

```

The results are show in @tbl-v1-2-12-comparison.

```{r}
#| label: tbl-v1-2-12-comparison
#| tbl-cap: "SomaHRD v1.2.8 vs 1.2.12 results"
#| echo: FALSE

knitr::kable(seqone_tab2_wide)

```

All outputs are identical except for the low_tumour_fraction and rad51b_cn.

```{r}
#| label: check-v-1-2-12
#| include: TRUE

identical(seqone_tab2_wide$lga_v1.2.8, seqone_tab2_wide$lga_v1.2.12)

identical(seqone_tab2_wide$lpc_v1.2.8, seqone_tab2_wide$lpc_v1.2.12)

identical(seqone_tab2_wide$score_v1.2.8, seqone_tab2_wide$score_v1.2.12)

identical(seqone_tab2_wide$status_v1.2.8, seqone_tab2_wide$status_v1.2.12)

identical(seqone_tab2_wide$brca_status_v1.2.8, seqone_tab2_wide$brca_status_v1.2.12)

identical(seqone_tab2_wide$brca_mutation_v1.2.8, seqone_tab2_wide$brca_mutation_v1.2.12)

identical(seqone_tab2_wide$ccne1_cn_v1.2.8, seqone_tab2_wide$ccne1_cn_v1.2.12)

identical(seqone_tab2_wide$rad51b_cn_v1.2.8, seqone_tab2_wide$rad51b_cn_v1.2.12)

identical(seqone_tab2_wide$coverage_v1.2.8, 
          seqone_tab2_wide$coverage_v1.2.12)

identical(seqone_tab2_wide$pct_mapped_reads_v1.2.8,
          seqone_tab2_wide$pct_mapped_reads_v1.2.12)

identical(seqone_tab2_wide$pct_tum_cell_v1.2.8,
          seqone_tab2_wide$pct_tum_cell_v1.2.12)

identical(seqone_tab2_wide$gi_confidence_v1.2.8,
          seqone_tab2_wide$gi_confidence_v1.2.12)

identical(seqone_tab2_wide$low_tumor_fraction_v1.2.8,
          seqone_tab2_wide$low_tumor_fraction_v1.2.12)

```

The difference for rad51b_cn was due to the value being formatted as a date, not
a number, in the original Excel.

```{r}
#| label: rad51b_cn
#| echo: FALSE

rad51b_tbl <- seqone_tab2_wide |> 
  filter(rad51b_cn_v1.2.8 != rad51b_cn_v1.2.12) |> 
  select(sample, rad51b_cn_v1.2.8, rad51b_cn_v1.2.12)

knitr::kable(rad51b_tbl)

```

```{r}
#| label: v1-2-12-low-tumour-fraction
#| echo: FALSE

low_tumour_fraction_tbl <- seqone_tab2_wide |> 
  filter(low_tumor_fraction_v1.2.8 != low_tumor_fraction_v1.2.12) |> 
  select(sample, low_tumor_fraction_v1.2.8, low_tumor_fraction_v1.2.12) |> 
  mutate(difference = abs(low_tumor_fraction_v1.2.8 - 
                            low_tumor_fraction_v1.2.12))

knitr::kable(low_tumour_fraction_tbl)

```

The differences for low_tumour_fraction were again very small, ranging from
`r min(low_tumour_fraction_tbl$difference)` to 
`r max(low_tumour_fraction_tbl$difference)`.

```{r}
#| label: low-tumour-fraction-full-comparison
#| echo: FALSE

low_tumour_v7_v12 <- seqone_tab1_wide |> 
              select(sample, low_tumor_fraction_v1.2.7) |> 
  left_join(seqone_tab2_wide |> 
              select(sample, low_tumor_fraction_v1.2.12),
            by = "sample") |> 
  mutate(difference = abs(low_tumor_fraction_v1.2.7 -
                          low_tumor_fraction_v1.2.12))

knitr::kable(low_tumour_v7_v12)

```

Comparing v1.2.7 to v1.2.12, the low tumour fraction varies from 
`r min(low_tumour_v7_v12$difference)` to 
`r max(low_tumour_v7_v12$difference)`.

The low tumour fraction metric is not used in the analysis of SeqOne data by
Manchester GLH scientists, and hence these minor changes are not expected to
impact the accuracy of the method.

# Verification of SomaHRD v1.2.13

On 2024-01-09 SeqOne informed us the SomaHRD version had updated again to 
v1.1.13. 

SeqOne sent analysed results for the 10 validation samples which are saved
here:

"S:/central shared/Genetics/Mol_Shared/Development.Team/SeqOne HRD/data/live_service/INC9891/somaHRD 1.2.12 vs 1.2.13.xlsx"

I checked every column in the results table - all were identical except for 
the low_tumour_fraction calculation.

```{r}
#| include: FALSE

seqone_v13_data <- read_excel(path = paste0(data_filepath,
                                         "live_service/INC9891/",
                                        "somaHRD 1.2.12 vs 1.2.13.xlsx"),
                           sheet = "Feuil1",
                           col_types = 
                             c("text", "date", "text", 
                               "numeric", "numeric", "numeric", 
                               "text", "text", "text",
                               "numeric", "numeric", "numeric", 
                               "numeric", "numeric", "numeric", 
                               "numeric")) |> 
  janitor::clean_names()

seqone_v13_wide <-  seqone_v13_data |> 
  pivot_wider(id_cols = sample,
              names_from = somahrd_version,
              values_from = -c(sample, somahrd_version))

identical(seqone_v13_wide$lga_v1.2.12, seqone_v13_wide$lga_v1.2.13)

identical(seqone_v13_wide$lpc_v1.2.12, seqone_v13_wide$lpc_v1.2.13)

identical(seqone_v13_wide$score_v1.2.12, seqone_v13_wide$score_v1.2.13)

identical(seqone_v13_wide$status_v1.2.12, seqone_v13_wide$status_v1.2.13)

identical(seqone_v13_wide$brca_status_v1.2.12, seqone_v13_wide$brca_status_v1.2.13)

identical(seqone_v13_wide$brca_mutation_v1.2.12, 
          seqone_v13_wide$brca_mutation_v1.2.13)

identical(seqone_v13_wide$ccne1_cn_v1.2.12,
          seqone_v13_wide$ccne1_cn_v1.2.13)

identical(seqone_v13_wide$rad51b_cn_v1.2.12,
          seqone_v13_wide$rad51b_cn_v1.2.13)

identical(seqone_v13_wide$coverage_v1.2.12, 
          seqone_v13_wide$coverage_v1.2.13)

identical(seqone_v13_wide$pct_mapped_reads_v1.2.12,
          seqone_v13_wide$pct_mapped_reads_v1.2.13)

identical(seqone_v13_wide$pct_tum_cell_v1.2.12,
          seqone_v13_wide$pct_tum_cell_v1.2.13)

identical(seqone_v13_wide$gi_confidence_v1.2.12,
          seqone_v13_wide$gi_confidence_v1.2.13)

identical(seqone_v13_wide$low_tumor_fraction_v1.2.12,
          seqone_v13_wide$low_tumor_fraction_v1.2.13)

```

```{r}
#| echo: FALSE

v13_low_tumour_table <- seqone_v13_wide |> 
  select(sample, low_tumor_fraction_v1.2.12,
         low_tumor_fraction_v1.2.13) |> 
  mutate(diff = abs(low_tumor_fraction_v1.2.12 - 
                      low_tumor_fraction_v1.2.13))

knitr::kable(v13_low_tumour_table)

```

The maximum difference was 
`r max(v13_low_tumour_table$diff)`.

In the same way as the previous verification, the variation in low_tumour_fraction
will not impact the accuracy of the SomaHRD method.

# Conclusion

SeqOne SomaHRD v1.2.13 is accepted for clinical use.
