---
title: "INC9096 Investigation"
author: "Joe Shaw"
format: html
date: today
date-format: "DD/MM/YYYY"
---

This is an investigation into worksheet WS140359. When analysed on the SeqOne pipeline on 01/04/2024 and 02/04/2024, all 5 samples failed the genomic instability confidence ("robsutness") metric with very low values, and two new columns (“gi_confidence_min” and “gi_confidence_max”) appeared in the exported csv files.
 
```{r}
#| label: packages-and-functions
#| include: false

library("tidyverse")
library("ggpubr")
library("readxl")
library("here")
library("patchwork")
library("rvest")

source(here::here("functions/hrd_functions.R"))

```

```{r}
#| label: WS140359-data
#| include: FALSE

ws140359_csvs <- list.files(here::here("data/WS140359 data/"),
                            full.names = TRUE,
                            pattern = ".csv")

ws140359_pdfs <- list.files(here::here("data/WS140359 data/"),
                            full.names = TRUE,
                            pattern = ".pdf")

ws140359_csv_data <- ws140359_csvs |> 
  map(\(ws140359_csvs) read_csv(ws140359_csvs,
                                 show_col_types = FALSE)) |> 
  list_rbind() |> 
  mutate(analysis_date = factor(analysis_date))

ws140359_pdf_data <- ws140359_pdfs |> 
  map(\(ws140359_pdfs) read_seqone_report(file = ws140359_pdfs,
                                          version = "1.2")) |> 
        list_rbind()

```

SeqOne addressed the issue and the samples were re-analysed on 04/04/2024: all samples now passed the genomic instability confidence threshold, and all other variables were the same.

```{r}
#| label: output-comparison
#| echo: FALSE

comparison_tbl <- ws140359_csv_data |> 
  select(-c(brca_status, brca_mutation, somahrd_version)) |> 
  arrange(sample, analysis_date)

knitr::kable(comparison_tbl)

```

## Versions

Looking at the HTML technical report files for a selection of samples shows that the versions of the components of the SomaHRD pipeline have changed since the validation.

```{r}
#| label: pipeline-versions
#| echo: FALSE

seqone_html_files <- list.files(path = here::here("data/html_reports/"), pattern = ".html", full.names = TRUE)

seqone_html_collated <- seqone_html_files |> 
  map(\(seqone_html_files) parse_seqone_html(seqone_html_files)) |> 
  list_rbind()

somahrd_pipeline_versions <- seqone_html_collated |> 
  filter(name == "SomaHRD pipeline") |> 
  arrange(worksheet)

hrd_prediction_versions <- seqone_html_collated |> 
  filter(name == "HRD prediction") |> 
  arrange(worksheet)

knitr::kable(somahrd_pipeline_versions)

knitr::kable(hrd_prediction_versions)

```
