---
title: "Robustness Confidence Intervals"
author: "Joe Shaw, Clinical Scientist (CS20980)"
date: today
date-format: "DD/MM/YYYY"
format: pdf
editor: visual
---
```{r}
#| label: packages
#| include: FALSE

library(tidyverse)

source(here::here("functions/hrd_functions.R"))

```
```{r}
#| label: load-collated-data
#| include: FALSE

# Data collated from PDFs in hrd_service_audit.qmd script.

all_hrd_data_pdf <- read.csv(here::here("data/seqone_collated_data/seqone_collated_pdf_data.csv"))

```
SomaHRD v1.3 includes confidence intervals for the confidence in genomic instability metric. If the confidence intervals span 0.5 then the sample is classified as inconclusive. The script for this analysis is on my [Github](https://github.com/joeshaw824/homologous_recombination_deficiency).

```{r}
#| label: add-confidence-intervals
#| include: FALSE

error_bar_qc_check_threshold <- 0.5

all_data_mod <- all_hrd_data_pdf |> 
  # Remove data from INC9096 (WS140359_1) and QiaSymphony validation worksheet (WS136827)
  filter(!worksheet %in% c("WS140359_1", "WS136827")) |> 
  mutate(
    std_dev = 1 - robustness,
    hrd_score_minus_sd = seqone_hrd_score - std_dev,
    hrd_score_plus_sd = seqone_hrd_score + std_dev,
    hrd_score_min = case_when(
      
      hrd_score_minus_sd < 0 ~0,
      
      hrd_score_minus_sd > 1 ~1,
      
      TRUE ~hrd_score_minus_sd),
    
    hrd_score_max = case_when(
      
      hrd_score_plus_sd < 0 ~0,
      
      hrd_score_plus_sd > 1 ~1,
      
      TRUE ~hrd_score_plus_sd),
    
    error_bar_width = hrd_score_max - hrd_score_min,
    
    error_bar_qc_check = case_when(
      
      hrd_score_max > error_bar_qc_check_threshold &
        
        hrd_score_min > error_bar_qc_check_threshold ~"pass",
      
      hrd_score_max < error_bar_qc_check_threshold &
        
        hrd_score_min < error_bar_qc_check_threshold ~"pass",
      
      hrd_score_max >= error_bar_qc_check_threshold &
        
        hrd_score_min <= error_bar_qc_check_threshold ~"fail"))

```

This table shows whether samples in our dataset would be considered "inconclusive" based on the confidence intervals overlapping 0.5. This is only for samples where the overall HRD status was not "NON-CONCLUSIVE".

```{r}
#| label: summary-table
#| echo: FALSE

summary_table <- all_data_mod |> 
  filter(seqone_hrd_status != "NON-CONCLUSIVE") |> 
  count(error_bar_qc_check) |> 
  mutate(percentage = round((n/sum(n)) * 100, 1))

knitr::kable(summary_table)

```

```{r}
#| label: plot-confidence-intervals
#| echo: FALSE

conf_int_plot <- all_data_mod |> 
  filter(seqone_hrd_status != "NON-CONCLUSIVE" &
           error_bar_qc_check == "fail") |> 
  ggplot(aes(x = reorder(sample_id, seqone_hrd_score),
                         y = seqone_hrd_score)) +
  geom_point(pch = 21, size = 3, aes(fill = seqone_hrd_status)) +
  scale_fill_manual(values = c(safe_blue, safe_red)) +
  geom_errorbar(aes(ymin = hrd_score_min, ymax = hrd_score_max)) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "Samples with error bars which cross 0.5 threshold",
       subtitle = "Only samples with conclusive results included",
       y = "SeqOne HRD Score", x = "") +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  ylim(0, 1)

conf_int_plot

```
